---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MR-SENSEMAKR

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/carloscinelli/mrsensemakr/branch/master/graph/badge.svg)](https://codecov.io/gh/carloscinelli/mrsensemakr?branch=master)
[![Travis build status](https://travis-ci.com/carloscinelli/mrsensemakr.svg?branch=master)](https://travis-ci.com/carloscinelli/mrsensemakr)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/carloscinelli/mrsensemakr?branch=master&svg=true)](https://ci.appveyor.com/project/carloscinelli/mrsensemakr)
<!-- badges: end -->

The R package `mrsensemakr` implements sensitivity analysis tools for Mendelian Randomization, as discussed in [Cinelli et al (2020). Robust Mendelian randomization in the presence of residual population stratification, batch effects and horizontal pleiotropy.](https://www.biorxiv.org/content/10.1101/2020.10.21.347773v1)

## Development version

<!-- commit -->

To install the development version on GitHub make sure you have the R package `devtools` installed. Also make sure to have the latest version of `sensemakr` [(link)](https://github.com/carloscinelli/sensemakr) installed.

``` r
# install.packages("devtools")
devtools::install_github("carloscinelli/sensemakr")
devtools::install_github("carloscinelli/mrsensemakr")
```

CRAN version coming soon.

## Basic usage



```{r example, dpi = 600}
## loads package
library(mrsensemakr)

## simulated data example
data("sim_data")

## create vectors indicating variable names in the data
outcome    <- "out.trait" # name of outcome trait
exposure   <- "exp.trait" # name of exposure trait
instrument <- "prs" # genetic instrument (e.g, polygenic risk score)
age.sex    <- c("age", "sex") # age and sex variables (if applicable)
alc.smok   <- c("alcohol", "smoking") # putative pleoitropic vars.
pcs        <- paste0("pc", 1:20) # first 20 principal components pc1 ... pc20

## runs MR sensitivity analysis
mr.sense <- mr_sensemakr(outcome = outcome,
                         exposure = exposure,
                         instrument = instrument,
                         covariates = c(age.sex, alc.smok, pcs), 
                         data = sim_data, 
                         benchmark_covariates = list(alc.smok = alc.smok,
                                                     pcs = pcs))
## print results
mr.sense


## sensitivity contour plots
plot(mr.sense, 
     benchmark_covariates = list(alc.smok = alc.smok,
                                 pcs = pcs),
     k = list(alc.smok = 1, pcs = 1))
```

## Simulations

Code to reproduce the simulations of the paper can be found in the vignettes folder.
